
```{r}
directory <- "/Users/maggiepeng/desktop/htseq"
sampleFiles <- list.files(directory)
sampleCultivar <- c("MSK","MSK","MSK","MSK","MSK","MTH","MTH","MTH")
samplePart <- c("leaf","root","aril","stem","aril","aril","aril","aril")
sampleTable <- data.frame(sampleName=sub(".count","",sampleFiles),fileName=sampleFiles, cultivar=sampleCultivar, part=samplePart)
sampleTable$cultivar <- factor(sampleTable$cultivar)
sampleTable$part <- factor(sampleTable$part)
```

1, Differential expression analysis among different cultivar in aril (Monthong VS MusangKing)
Load sample data from HTSeq files
```{r}
library("DESeq2")
sampleTable_aril <- sampleTable[which(sampleTable$part=="aril"),]
sampleTable_aril$cultivar <- relevel(sampleTable_aril$cultivar, ref ="MTH")
ddsCultivar <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable_aril, directory = directory, design= ~ cultivar)
ddsCultivar
```
Differential expression analysis
```{r}
ddsCultivar <- DESeq(ddsCultivar)
resCultivar <- results(ddsCultivar)
resCultivar[order(abs(resCultivar$log2FoldChange), decreasing= T), ]
summary(resCultivar)
```
visualization of results
```{r}
DESeq2::plotMA(resCultivar,ylim=c(-10,10))

#Filter significant result and make a new fig
library(dplyr)
resCultivar1 = as.data.frame(resCultivar)
resCultivar1 = mutate(resCultivar1, sig=ifelse(resCultivar1$padj<0.1, "FDR<0.1", "Not Sig"))
resCultivar1[which(abs(resCultivar1$log2FoldChange)<1.0), "sig"]="Not Sig"

library(ggplot2)
ggplot(resCultivar1, aes(log2FoldChange, -log10(padj)))+geom_point(aes(col=sig))+scale_color_manual(values = c("red","black"))
```

2.1, Differential expression analysis among different parts in Musang King (stem vs aril)
```{r}
sampleTable_MK <- sampleTable[which(sampleTable$cultivar=="MSK"&(sampleTable$part=="stem"|sampleTable$part=="aril")),]
sampleTable_MK
sampleTable_MK$part <- relevel(sampleTable_MK$part, ref ="stem")
ddsPart <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable_MK, directory = directory, design= ~ part)
ddsPart
```
Differential analysis
```{r}
ddsPart <- DESeq(ddsPart)
resPart <- results(ddsPart)
resPart[order(resPart$log2FoldChange, decreasing= T), ]
summary(resPart)
```
visualization of results
```{r}
DESeq2::plotMA(resPart,ylim=c(-10,10))

#Filter significant result and make a new fig
resPart1 = as.data.frame(resPart)
resPart1 = mutate(resPart1, sig=ifelse(resPart1$padj<0.1, "FDR<0.1", "Not Sig"))
resPart1[which(abs(resPart1$log2FoldChange)<1.0), "sig"]="Not Sig"
ggplot(resPart1, aes(log2FoldChange, -log10(padj)))+geom_point(aes(col=sig))+scale_color_manual(values = c("red","black"))
```

2.2, Differential expression analysis among different parts in Musang King (leaf vs aril)
```{r}
#extract subset from sampleTable
sampleTable_MK_LA <- sampleTable[which(sampleTable$cultivar=="MSK"&(sampleTable$part=="leaf"|sampleTable$part=="aril")),]
sampleTable_MK_LA$part <- relevel(sampleTable_MK_LA$part, ref ="leaf")
ddsPart_LA <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable_MK_LA, directory = directory, design= ~ part)
ddsPart_LA
```
Differential analysis
```{r}
ddsPart_LA <- DESeq(ddsPart_LA)
resPart_LA <- results(ddsPart_LA)
resPart_LA[order(resPart_LA$log2FoldChange, decreasing= T), ]
summary(resPart_LA)
```
visualization of results
```{r}
DESeq2::plotMA(resPart_LA,ylim=c(-10,10))

#Filter significant result and make a new fig
resPart_LA1 = as.data.frame(resPart_LA)
resPart_LA1 = mutate(resPart_LA1, sig=ifelse(resPart_LA1$padj<0.1, "FDR<0.1", "Not Sig"))
resPart_LA1[which(abs(resPart_LA1$log2FoldChange)<1.0), "sig"]="Not Sig"
ggplot(resPart_LA1, aes(log2FoldChange, -log10(padj)))+geom_point(aes(col=sig))+scale_color_manual(values = c("red","black"))
```

2.3, Differential expression analysis among different parts in Musang King (root vs aril)
```{r}
#extract subset from sampleTable
sampleTable_MK_RA <- sampleTable[which(sampleTable$cultivar=="MSK"&(sampleTable$part=="root"|sampleTable$part=="aril")),]
sampleTable_MK_RA$part <- relevel(sampleTable_MK_RA$part, ref ="root")
ddsPart_RA <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable_MK_RA, directory = directory, design= ~ part)
ddsPart_RA
```

Differential analysis
```{r}
ddsPart_RA <- DESeq(ddsPart_RA)
resPart_RA <- results(ddsPart_RA)
resPart_RA[order(resPart_RA$log2FoldChange, decreasing= T), ]
summary(resPart_RA)
```

visualization of results
```{r}
DESeq2::plotMA(resPart_RA,ylim=c(-10,10))

#Filter significant result and make a new fig
resPart_RA1 = as.data.frame(resPart_RA)
resPart_RA1 = mutate(resPart_RA1, sig=ifelse(resPart_RA1$padj<0.1, "FDR<0.1", "Not Sig"))
resPart_RA1[which(abs(resPart_RA1$log2FoldChange)<1.0), "sig"]="Not Sig"
ggplot(resPart_RA1, aes(log2FoldChange, -log10(padj)))+geom_point(aes(col=sig))+scale_color_manual(values = c("red","black"))
```


